---
title: "PA1_template.Rmd"
author: "Sebastian Diaz de Valdes"
date: "15-04-2020"
output: html_document
---
Project 1
================
**Loading and preprocessing the data**
```{R, echo=TRUE}
data<-read.csv("activity.csv" )
head(data)
```

**What is mean total number of steps taken per day and the median?**

*-Calculate and report the mean and median total number of steps taken per day*
```{R, echo=TRUE}
library(dplyr)
newdata <- na.omit(data)
bydays<-list(newdata$date)
sumsbyday<-aggregate(newdata[c("steps")], by=bydays,sum)
colnames(sumsbyday)[colnames(sumsbyday) == 'Group.1'] <- 'Date'
colnames(sumsbyday)[colnames(sumsbyday) == 'steps'] <- 'Total_Steps'
mean(sumsbyday$Total_Steps)
median((sumsbyday$Total_Steps))
```

*-Make a histogram of the total number of steps taken each day*
```{R, echo=TRUE}
hist(x = sumsbyday$Total_Steps,breaks=30,xlab = "Total Number of Steps per Day", main = "Histogram of Total Number of Steps per Day")
```


**What is the average daily activity pattern?**

*-which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?*
```{r, echo=TRUE}
intervals<-aggregate(steps ~ interval, data=data, mean)
intervalMax<-subset(intervals,intervals$steps==(max(intervals$steps)))
maxi<-intervalMax[1,1]
maxi2<-intervalMax[1,2]
```

The max 5-minutes interval is `r maxi` and its values is `r maxi2`

*-Make a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)*

```{R, echo=TRUE}
library(ggplot2)
library(dplyr)
 ggplot(intervals, aes(x=interval, y=steps)) +
     geom_line() + 
     labs(x = "Intervals", y = "Steps",title ="Average daily activity pattern")
```


**Imputing missing values**

*-Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)*
```{r,echo=TRUE}
missingvalues<-sum(is.na(data))
missingvalues
colSums(is.na(data))
```


```{r,echo=TRUE}
mean(data$steps, na.rm = T)
```

*-Create a new dataset that is equal to the original dataset but with the missing data filled in.*
```{r,echo=TRUE}
dataimpute <- data
dataimpute$steps[is.na(dataimpute$steps)] <- mean(dataimpute$steps, na.rm = T)
colSums(is.na(dataimpute))
```

```{r,echo=TRUE}
dataimpute$date<-as.Date(dataimpute$date)
sumsbyday2<-aggregate(steps~date, data=dataimpute,sum)
colnames(sumsbyday2)[colnames(sumsbyday2) == 'steps'] <- 'Total_Steps'
head(sumsbyday2)
```

*-Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?*



```{r,echo=TRUE}
hist(x = sumsbyday2$Total_Steps,breaks=30,xlab = "Total Number of Steps per Day", main = "Histogram of Total Number of Steps per Day with NA replaced")
```

*- Calculate and report the mean and median total number of steps taken per day*
```{r,echo=TRUE}
mean(sumsbyday2$Total_Steps)
median((sumsbyday2$Total_Steps))
```
The mean and mediam does not change!!


**Are there differences in activity patterns between weekdays and weekends?**


```{r,echo=TRUE}
dataimpute$weekdays <- weekdays(dataimpute$date)
dataimpute$weeks[(dataimpute$weekdays == "sabado" | dataimpute$weekdays == "domingo")] <- "weekend"
dataimpute$weeks[!(dataimpute$weekdays == "sabado" | dataimpute$weekdays == "domingo")] <- "weekdays"
library(plyr)
finaldf<- ddply(dataimpute, c("interval", "weeks"), function(x) apply(x[1],2, mean))
head(finaldf)
```


```{r,echo=TRUE}
library(ggplot2)
library(dplyr)
ggplot(finaldf, aes(x=interval, y=steps)) +
     geom_line() + 
     xlab("") +
     labs(x = "Intervals", y = "Steps",title ="Average daily activity pattern" )+
     facet_wrap(~weeks)
```

We can observe the differences!
